### 2020-12-31 paper/model.tex

\documentclass[border=8pt, multi, tikz]{standalone} 
\usepackage{import}
\subimport{./layers/}{init}
\usetikzlibrary{positioning}
\usetikzlibrary{3d} %for including external image

\def\ConvColor{rgb:yellow,5;red,2.5;white,5}
\def\ConvReluColor{rgb:yellow,5;red,5;white,5}
\def\PoolColor{rgb:red,1;black,0.3}
\def\UnpoolColor{rgb:blue,2;green,1;black,0.3}
\def\FcColor{rgb:blue,5;red,2.5;white,5}
\def\FcReluColor{rgb:blue,5;red,5;white,4}
\def\SoftmaxColor{rgb:magenta,5;black,7}   
\def\SumColor{rgb:blue,5;green,15}

\newcommand{\copymidarrow}{\tikz \draw[-Stealth,line width=0.8mm,draw={rgb:blue,4;red,1;green,1;black,3}] (-0.3,0) -- ++(0.3,0);}

\begin{document}
\begin{tikzpicture}
\tikzstyle{connection}=[ultra thick,every node/.style={sloped,allow upside down},draw=\edgecolor,opacity=0.7]
\tikzstyle{copyconnection}=[ultra thick,every node/.style={sloped,allow upside down},draw={rgb:blue,4;red,1;green,1;black,3},opacity=0.7]

\node[canvas is zy plane at x=0] (temp) at (0,0,0) {\includegraphics[width=4cm,height=4cm]{./sample.png}};

\pic[shift={(1,0,0)}] at (0,0,0) 
    {Box={
        name=conv1,
        caption= ,
        xlabel={{32, }},
        zlabel=64,
        fill=\ConvColor,
        height=13.5,
        width=9.0,
        depth=13.5
        }
    };

\pic[shift={(1,0,0)}] at (conv1-east) 
    {Box={
        name=conv2,
        caption= ,
        xlabel={{32, }},
        zlabel=64,
        fill=\ConvColor,
        height=13.5,
        width=9.0,
        depth=13.5
        }
    };

\pic[shift={ (1,0,0) }] at (conv2-east) 
    {Box={
        name=pool1,
        caption= ,
        fill=\PoolColor,
        opacity=0.5,
        height=9.0,
        width=9.0,
        depth=9.0
        }
    };

\pic[shift={(1,0,0)}] at (pool1-east) 
    {Box={
        name=conv3,
        caption= ,
        xlabel={{64, }},
        zlabel=32,
        fill=\ConvColor,
        height=9.0,
        width=13.5,
        depth=9.0
        }
    };

\pic[shift={ (1,0,0) }] at (conv3-east) 
    {Box={
        name=pool2,
        caption= ,
        fill=\PoolColor,
        opacity=0.5,
        height=6.0,
        width=13.5,
        depth=6.0
        }
    };

\pic[shift={(1,0,0)}] at (pool2-east) 
    {Box={
        name=conv4,
        caption= ,
        xlabel={{128, }},
        zlabel=16,
        fill=\ConvColor,
        height=6.0,
        width=14.850000000000001,
        depth=6.0
        }
    };

\pic[shift={ (1,0,0) }] at (conv4-east) 
    {Box={
        name=pool3,
        caption= ,
        fill=\PoolColor,
        opacity=0.5,
        height=4,
        width=14.850000000000001,
        depth=4
        }
    };

\pic[shift={(1,0,0)}] at (pool3-east) 
    {Box={
        name=conv5,
        caption= ,
        xlabel={{256, }},
        zlabel=8,
        fill=\ConvColor,
        height=4,
        width=16.335000000000004,
        depth=4
        }
    };

\pic[shift={ (1,0,0) }] at (conv5-east) 
    {Box={
        name=pool4,
        caption= ,
        fill=\PoolColor,
        opacity=0.5,
        height=2,
        width=16.335000000000004,
        depth=2
        }
    };

\pic[shift={(1,0,0)}] at (pool4-east) 
    {Box={
        name=conv6,
        caption= ,
        xlabel={{, }},
        zlabel=2048,
        fill=\ConvColor,
        height=32.67000000000001,
        width=1,
        depth=32.67000000000001
        }
    };

\pic[shift={(1,0,0)}] at (conv6-east) 
    {Box={
        name=conv7,
        caption= ,
        xlabel={{, }},
        zlabel=2048,
        fill=\ConvColor,
        height=32.67000000000001,
        width=1,
        depth=32.67000000000001
        }
    };

\pic[shift={(1,0,0)}] at (conv7-east) 
    {Box={
        name=softmax,
        caption= ,
        xlabel={{" ","dummy"}},
        zlabel=8,
        fill=\SoftmaxColor,
        opacity=0.8,
        height=3,
        width=1.5,
        depth=25
        }
    };

\draw [connection]  (input-east)    -- node {\midarrow} (conv1-west);

\draw [connection]  (conv1-east)    -- node {\midarrow} (conv2-west);

\draw [connection]  (conv2-east)    -- node {\midarrow} (pool1-west);

\draw [connection]  (pool1-east)    -- node {\midarrow} (conv3-west);

\draw [connection]  (conv3-east)    -- node {\midarrow} (pool2-west);

\draw [connection]  (pool2-east)    -- node {\midarrow} (conv4-west);

\draw [connection]  (conv4-east)    -- node {\midarrow} (pool3-west);

\draw [connection]  (pool3-east)    -- node {\midarrow} (conv5-west);

\draw [connection]  (conv5-east)    -- node {\midarrow} (pool4-west);

\draw [connection]  (pool4-east)    -- node {\midarrow} (conv6-west);

\draw [connection]  (conv6-east)    -- node {\midarrow} (conv7-west);

\draw [connection]  (conv7-east)    -- node {\midarrow} (softmax-west);

\end{tikzpicture}
\end{document}